\section{Conclusion}

In this paper we present an automated design flow for design space
exploration of dataflow designs based on a novel language, MaxC, and
aspect-driven compilation. We show that this approach meets the
initial performance, portability, integration and productivity
requirements. We implement a high performance dataflow design for RTM
to show that our approach can be used to achieve significant speedups
of up to 100X over software only versions by exploiting the run-time
reconfiguration capabilities of FPGAs. The automated optimisation
process driven by exploration and implementation aspects improves
productivity and design portability, allowing optimisations strategies
to be automatically adapted to various platforms. Compilation aspects
can be used to identify design configurations that improve compilation
speed, further speeding up the development process. Finally we show
that MaxC is a concise language (based on the LOC count) and that
maintaining compatibility with C99 simplifies the translation of
existing applications.

Future work includes extending our approach by introducing new system
aspects that guide the translation and optimisation process from
existing C applications to dataflow designs. This includes the
development of system aspects that can be used to explore structural
transformations of the original application that improve parallelism
and computation versus communication ratio. The introduction of these
aspects could lead to a transparent design flow, that does not require
manual development of dataflow designs, greatly simplifying the
translation of existing software applications to high-performance
custom designs.
